Many-core applications to online track reconstruction in HEP experiments

### 2013-10-03 papers/2013/CHEP/paper.tex
boilerplate text, some placeholder figures

\documentclass[letterpaper]{jpconf}
\usepackage{graphicx}
\usepackage{iopams}
\usepackage[pdftex,
      colorlinks=true,
      urlcolor=blue,       % \href{...}{...} external (URL)
      filecolor=green,     % \href{...} local file
      linkcolor=red,       % \ref{...} and \pageref{...}
      pagebackref,
      pdfpagemode=UseNone,
      bookmarksopen=true]{hyperref}

@@

We showed initial studies on GPU performance in low-latency
environments ($\sim$100\,$\mu$s) in a previous paper~\cite{TIPP2011,
  NSS2012}.  In this paper we extend those studies measuring the
timing performance of different GPU cards in different data I/O
environments. The algorithm run on the GPU is a complete version of
the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
system at CDF~\cite{SVT1}.

The goal of this study is to investigate the strengths and weaknesses
of GPUs when applied in a low latency environment, with particular
emphasis on the data transfer latency to/from the GPU and the
algorithm latency for processing on the GPU in a manner similar to a
typical HEP trigger application.

@@

>>>>

\documentclass[letterpaper]{jpconf}
\usepackage{graphicx}
\usepackage{iopams}
\usepackage[pdftex,
      colorlinks=true,
      urlcolor=blue,       % \href{...}{...} external (URL)
      filecolor=green,     % \href{...} local file
      linkcolor=blue,       % \ref{...} and \pageref{...}
      pagebackref,
      pdfpagemode=UseNone,
      bookmarksopen=true]{hyperref}

@@

The ultimate goal of this study is to investigate the strengths and weaknesses
of GPUs when applied in a low latency environment, with particular
emphasis on the data transfer latency to/from the GPU and the
algorithm latency for processing on the GPU in a manner similar to a
typical HEP trigger application.

We showed initial studies on GPU performance in low-latency
environments ($\sim$100\,$\mu$s) in a previous paper~\cite{TIPP2011,
  NSS2012}.  In this paper we extend those studies measuring the
timing performance of different GPU cards in different data I/O
environments. The algorithm run on the GPU is a complete version of
the fast track fitting algorithm of the Silicon Vertex Tracker (SVT)
system at CDF~\cite{SVT1}. We also compare the performance of GPU
cards from different manufacturers (AMD vs NVIDIA) and different
computing environments (\textsc{cuda} and \textsc{opencl}) on the same
card.

One of the goals of the studies is to explore using a concrete
algorithm what gains one can achieve with varying degrees of changes
to the existing algorithm. Starting with a serial algorithm
implemented on a CPU, we test an ``embarrassingly parallel'' algorithm
on the Intel MIC environment. In this case each event is handled
independently by a core on the accelerator, and the parallelization is
achieved with only minor changes to the legacy code and is only
possible in the Intel MIC environment. Next we consider an algorithm
where we unroll three internal nested loops and run these in parallel
on a GPU, using the \textsc{cuda} environment. This second approach is
programmatically more complicated and less trivial to implement. In
neither case have we re-thought the basic algorithms or the data
structures used. To achieve optimal performance, these steps would
have to be taken.  As one might expect, the improvement from the first
approach is rather modest, and the second approach shows larger
performance gains.

@@

\section{SVT track fitting algorithm}

\section{Results}
\subsection{Data processing}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}

  \caption{Algorithm-only comparison for timing as a function of the
    number of track fits. We compare timing on CPUs (serial), Xeon Phi
    (embarrassingly parallel), and GPUs (fully parallel.) Conclusion
    is ...}
  \label{fig:algo_only_timing}
\end{figure}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}

  \caption{Speed-ups for algorithm-only comparison for timing as a
    function of the number of track fits. We compare timing on CPUs
    (serial), Xeon Phi (partially parallel), and GPUs (fully
    parallel.) We conclude ....}
  \label{fig:algo_only_speedup}
\end{figure}
In Fig.~\ref{fig:algo_only_timing} we compare the algorithm time as a
function of the number of fits performed, for the serial,
embarrassingly parallel and parallel algorithms. We see that the
embarrassingly parallel algorithm gives a modest increase with respect
to the serial (CPU) algorithm. Switching to a fully parallel algorithm
affords a much more significant speed
improvement. Fig.~\ref{fig:algo_only_speedup} shows the speed-up for
each... do we want to show this. Not a lot of new information.

\subsection{Breakdown of computing time}
\begin{figure}[tbp]
  \centering
  \includegraphics[width=0.9\linewidth]{../../2012/figures/DT_MC_SLINK_IB_APE_low}
  \caption{breakdown of computing time}
  \label{fig:breakdown}
\end{figure}
In Fig.~\ref{fig:breakdown} we show the fractional time spent in
various parts of the algorithm for the serial algorithm (on a CPU),
the embarrassingly parallel algorithm (on Intel MIC) and the parallel
algorithm (on NVidia GPUs), as a function of the number of
fits. Interesting features are ...
